---
title: "Additional marginal means tables"
author: "Quentin D. Read"
date: "3/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This document has some additional estimates of marginal means in table and figure form, as requested by the in house reviewer.

The reviewer requested the following:

### Model 1 (genotype)

-	Means separation of AUDPC for planting date
-	Means separation of AUDPC for genotype
-	Means separation of yield for planting date
-	Means separation of yield for irrigation 
-	Means separation of yield for genotype
-	Means separation of yield for PD x genotype interaction

### Model 2 (resistance)

-	~~Means separation of AUDPC for planting date~~ (same as model 1)
-	Means separation of AUDPC for resistance
-	~~Means separation of yield for planting date~~ (same as model 1)
-	~~Means separation of yield for irrigation~~ (same as model 1)
-	Means separation of yield for resistance type
-	Means separation of yield for PD x resistance type
- Means separation of yield for irrigation x resistance interaction (added 3/17)
- Means separation of AUDPC for irrigation x resistance interaction (added 3/17)

## Setup data and models

This is the same as in the other document.

```{r}
library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(multcomp)

windowsFonts(`fgbook` = windowsFont("Franklin Gothic Book"))
th <- theme_bw(base_family = 'fgbook',
               base_size = 14) + 
  theme(panel.grid = element_blank(),
        strip.background = element_blank())
theme_set(th)

dat <- fread('Datacom.csv', na.strings = '.', stringsAsFactors = TRUE)
dat[, PD := ordered(PD, labels = c('April','May','June'))]
dat[, MG := factor(MG)]
dat[, Rep := factor(Rep)]
dat[, Irr := factor(Irr, levels = c('Nonirr', 'Irr'))]
dat[, resistance := factor(ifelse(Strains %in% c("60561121", "60561331", "DS880", "DT974290"), 'R', 'S'))]

geno_names <- c("DT974290", "DS880", "60561121", "60561331", "LS980358", "Pharaoh", "6070242",  "DS97949")
geno_fullnames <- c('DT97-4290','DS-880','DS6056-1-1-2-1','6056-1-3-3-1','LS98-0358','Pharaoh','DS6070-2-4-2','DS-97949')
geno_lookup <- data.table(name = geno_names, full_name = geno_fullnames)

dat[, Strains := factor(Strains, levels = levels(Strains), labels = geno_lookup$full_name[match(levels(Strains), geno_lookup$name)])]
```

Fit model 1, with genotype main effects.

```{r}
lmm_audpc_bystrain <- lmer(log1p(AUDPC) ~ PD * Irr * Strains + (1|YR/Rep), data = dat)
lmm_yield_bystrain <- lmer(YldKgHa ~ PD * Irr * Strains + (1|YR/Rep), data = dat)
```

Fit model 2, with resistance group main effect.

```{r}
lmm_full_audpc <- lmer(log1p(AUDPC) ~ PD * Irr * resistance + (1|YR/Rep) + (1|Strains), data = dat)
lmm_full_yield <- lmer(YldKgHa ~ PD * Irr * resistance + (1|YR/Rep) + (1|Strains), data = dat)
```

## Estimate marginal means

Marginal means requested for model 1.

```{r}
emmeans1_audpc_pd <- emmeans(lmm_audpc_bystrain, ~ PD, mode = 'kenward-roger')
emmeans1_audpc_genotype <- emmeans(lmm_audpc_bystrain, ~ Strains, mode = 'kenward-roger')
emmeans1_yld_pd <- emmeans(lmm_yield_bystrain, ~ PD, mode = 'kenward-roger')
emmeans1_yld_irr <- emmeans(lmm_yield_bystrain, ~ Irr, mode = 'kenward-roger')
emmeans1_yld_genotype <- emmeans(lmm_yield_bystrain, ~ Strains, mode = 'kenward-roger')
emmeans1_yld_pdxgeno <- emmeans(lmm_yield_bystrain, ~ PD + Strains, mode = 'kenward-roger')

cld1_audpc_pd <- cld(emmeans1_audpc_pd, adjust = 'sidak', Letters = letters)
cld1_audpc_genotype <- cld(emmeans1_audpc_genotype, adjust = 'sidak', Letters = letters)
cld1_yld_pd <- cld(emmeans1_yld_pd, adjust = 'sidak', Letters = letters)
cld1_yld_irr <- cld(emmeans1_yld_irr, adjust = 'sidak', Letters = letters)
cld1_yld_genotype <- cld(emmeans1_yld_genotype, adjust = 'sidak', Letters = letters)
cld1_yld_pdxgeno <- cld(emmeans1_yld_pdxgeno, adjust = 'sidak', Letters = letters)

cols_to_transform <- c('emmean', 'lower.CL', 'upper.CL')

cld1_audpc_pd[, cols_to_transform] <- expm1(cld1_audpc_pd[, cols_to_transform])
cld1_audpc_genotype[, cols_to_transform] <- expm1(cld1_audpc_genotype[, cols_to_transform])
```

Marginal means requested for model 2.

```{r}
emmeans2_audpc_pd <- emmeans(lmm_full_audpc, ~ PD, mode = 'kenward-roger')
emmeans2_audpc_resistance <- emmeans(lmm_full_audpc, ~ resistance, mode = 'kenward-roger')
emmeans2_yld_pd <- emmeans(lmm_full_yield, ~ PD, mode = 'kenward-roger')
emmeans2_yld_irr <- emmeans(lmm_full_yield, ~ Irr, mode = 'kenward-roger')
emmeans2_yld_resistance <- emmeans(lmm_full_yield, ~ resistance, mode = 'kenward-roger')
emmeans2_yld_pdxresistance <- emmeans(lmm_full_yield, ~ PD + resistance, mode = 'kenward-roger')
emmeans2_yld_irrxresistance <- emmeans(lmm_full_yield, ~ Irr + resistance, mode = 'kenward-roger')
emmeans2_audpc_irrxresistance <- emmeans(lmm_full_audpc, ~ Irr + resistance, mode = 'kenward-roger')

cld2_audpc_pd <- cld(emmeans2_audpc_pd, adjust = 'sidak', Letters = letters)
cld2_audpc_resistance <- cld(emmeans2_audpc_resistance, adjust = 'sidak', Letters = letters)
cld2_yld_pd <- cld(emmeans2_yld_pd, adjust = 'sidak', Letters = letters)
cld2_yld_irr <- cld(emmeans2_yld_irr, adjust = 'sidak', Letters = letters)
cld2_yld_resistance <- cld(emmeans2_yld_resistance, adjust = 'sidak', Letters = letters)
cld2_yld_pdxresistance <- cld(emmeans2_yld_pdxresistance, adjust = 'sidak', Letters = letters)
cld2_yld_irrxresistance <- cld(emmeans2_yld_irrxresistance, adjust = 'sidak', Letters = letters)
cld2_audpc_irrxresistance <- cld(emmeans2_audpc_irrxresistance, adjust = 'sidak', Letters = letters)

cld2_audpc_pd[, cols_to_transform] <- expm1(cld2_audpc_pd[, cols_to_transform])
cld2_audpc_resistance[, cols_to_transform] <- expm1(cld2_audpc_resistance[, cols_to_transform])
cld2_audpc_irrxresistance[, cols_to_transform] <- expm1(cld2_audpc_irrxresistance[, cols_to_transform])
```

## Marginal means in table form

```{r, echo = FALSE}
var_lookup <- data.frame(short= c('PD','resistance','Irr','Strains'), long = c('planting date', 'resistance', 'irrigation', 'genotype'))

cld2kable <- function(clddat, include_vars, table_name) {
  clddat <- setDT(copy(as.data.frame(clddat)))
  if ('resistance' %in% include_vars) clddat[, resistance := factor(resistance, levels = c('R','S'), labels = c('resistant', 'susceptible'))]
  if ('Irr' %in% include_vars) clddat[, Irr := factor(Irr, levels = c('Irr', 'Nonirr'), labels = c('irrigated', 'non-irrigated'))]
  clddat[, .group := gsub(' ', '', .group)]
  clddat[, c('SE', 'df') := NULL]
  
  col_labels <- c(var_lookup$long[match(include_vars, var_lookup$short)], 'marginal mean', 'lower 95%', 'upper 95%', '')
  
  setorderv(clddat, cols = include_vars)
  
  knitr::kable(clddat,
               digits = 1,
               col.names = col_labels,
               caption = table_name)
}
```

### Model 1 (by genotype)

```{r, echo = FALSE}
cld2kable(cld1_audpc_pd, c('PD'), 'Marginal means table: AUDPC by planting date')
cld2kable(cld1_audpc_genotype, c('Strains'), 'Marginal means table: AUDPC by genotype')
cld2kable(cld1_yld_pd, c('PD'), 'Marginal means table: Yield by planting date')
cld2kable(cld1_yld_irr, c('Irr'), 'Marginal means table: Yield by irrigation')
cld2kable(cld1_yld_genotype, c('Strains'), 'Marginal means table: Yield by genotype')
cld2kable(cld1_yld_pdxgeno, c('PD', 'Strains'), 'Marginal means table: Yield by planting date &times; genotype')
```


### Model 2 (by resistance group)

The ones that the reviewer marked in blue that are almost exactly the same between Model 1 and Model 2 are not included here.

```{r, echo = FALSE}
cld2kable(cld2_audpc_resistance, c('resistance'), 'Marginal means table: AUDPC by resistance')
cld2kable(cld2_yld_resistance, c('resistance'), 'Marginal means table: Yield by resistance')
cld2kable(cld2_yld_pdxresistance, c('PD', 'resistance'), 'Marginal means table: Yield by planting date &times; resistance')
cld2kable(cld2_yld_irrxresistance, c('Irr', 'resistance'), 'Marginal means table: Yield by irrigation &times; resistance')
cld2kable(cld2_audpc_irrxresistance, c('Irr', 'resistance'), 'Marginal means table: AUDPC by irrigation &times; resistance')
```

## Marginal means in figure form

### Model 1 (by genotype)

```{r, echo = FALSE}
facet_irr <- facet_wrap(~ irr_toplot)
xaxis_pd <- scale_x_discrete(name = 'Planting Date', labels = c('April','May','June'))

colors_res <- scale_color_manual(values = c('forestgreen', 'goldenrod'), labels = c('resistant', 'susceptible'))
colors_pd <- scale_color_brewer(palette = 'Dark2')
shape_mg <- scale_shape_discrete(name = 'maturity group', labels = c('IV', 'V'))
grid_pd_irr <- facet_grid(PD ~ Irr, labeller = labeller(PD = c('1' = 'April', '2' = 'May', '3' = 'June')))
```

```{r plots by strain, echo = FALSE}
ggplot(cld1_audpc_pd, aes(x = PD, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  xaxis_pd +
  scale_y_log10(name = 'AUDPC', labels = scales::comma) +
  theme(legend.position = 'none')

cld1_audpc_genotype$Strains <- factor(cld1_audpc_genotype$Strains, levels = cld1_audpc_genotype$Strains[order(cld1_audpc_genotype$emmean)])

ggplot(cld1_audpc_genotype, aes(x = Strains, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  scale_y_log10(name = 'AUDPC', labels = scales::comma) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Genotypes')

ggplot(cld1_yld_pd, aes(x = PD, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  xaxis_pd +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  theme(legend.position = 'none')

cld1_yld_irr$Irr <- factor(cld1_yld_irr$Irr, levels = c('Irr', 'Nonirr'), labels = c('irrigated', 'non-irrigated'))

ggplot(cld1_yld_irr, aes(x = Irr, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  theme(legend.position = 'none', axis.title.x = element_blank())

# Arrange all genotypes with the same order
gen_order <- as.character(cld1_yld_genotype$Strains[order(cld1_yld_genotype$emmean)])

cld1_yld_genotype$Strains <- factor(cld1_yld_genotype$Strains, levels = gen_order)

ggplot(cld1_yld_genotype, aes(x = Strains, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Genotypes')

cld1_yld_pdxgeno$Strains <- factor(cld1_yld_pdxgeno$Strains, levels = gen_order)

ggplot(cld1_yld_pdxgeno, aes(x = Strains, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Genotypes') +
  facet_wrap(~ PD, nrow = 3)
```

### Model 2 (by resistance group)

Again, the ones marked in blue by the reviewer that will be essentially identical to Model 1 are not included here.

```{r plots by resistance, echo = FALSE}
ggplot(cld2_audpc_resistance, aes(x = resistance, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  scale_y_log10(name = 'AUDPC', labels = scales::comma) +
  scale_x_discrete(labels = c('resistant', 'susceptible')) +
  theme(legend.position = 'none',
        axis.title.x = element_blank())

ggplot(cld2_yld_resistance, aes(x = resistance, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75) +
  geom_point(size = 2) + 
  geom_text(nudge_x = 0.15, hjust = 0) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  scale_x_discrete(labels = c('resistant', 'susceptible')) +
  theme(legend.position = 'none',
        axis.title.x = element_blank())

pd <- position_dodge(width = 0.5)

ggplot(cld2_yld_pdxresistance, aes(x = PD, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group), group = interaction(PD, resistance))) +
  geom_errorbar(width = 0.15, size = 0.75, aes(color = resistance), position = pd) +
  geom_point(size = 2, position = pd) + 
  geom_text(hjust = -.5, position = pd) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  xaxis_pd +
  theme(legend.position = c(0.8, 0.15), legend.title = element_blank()) +
  colors_res 

cld2_yld_irrxresistance$Irr <- factor(cld2_yld_irrxresistance$Irr, levels = c('Irr', 'Nonirr'), labels = c('irrigated', 'non-irrigated'))

ggplot(cld2_yld_irrxresistance, aes(x = Irr, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group), group = interaction(Irr, resistance))) +
  geom_errorbar(width = 0.15, size = 0.75, aes(color = resistance), position = pd) +
  geom_point(size = 2, position = pd) + 
  geom_text(hjust = -.5, position = pd) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma) +
  theme(legend.position = c(0.8, 0.9), legend.title = element_blank(), axis.title.x = element_blank()) +
  colors_res 

cld2_audpc_irrxresistance$Irr <- factor(cld2_audpc_irrxresistance$Irr, levels = c('Irr', 'Nonirr'), labels = c('irrigated', 'non-irrigated'))

ggplot(cld2_audpc_irrxresistance, aes(x = Irr, y = emmean, ymin = lower.CL, ymax = upper.CL, label = gsub(' ', '', .group), group = interaction(Irr, resistance))) +
  geom_errorbar(width = 0.15, size = 0.75, aes(color = resistance), position = pd) +
  geom_point(size = 2, position = pd) + 
  geom_text(hjust = -.5, position = pd) +
  scale_y_log10(name = 'AUDPC', labels = scales::comma) +
  theme(legend.position = c(0.85, 0.15), legend.title = element_blank(), axis.title.x = element_blank()) +
  colors_res 
```

### Additional figure

Figure added on 24 March. This figure includes marginal means of yield for all combinations of genotype, planting date, and irrigation. Colors of the bars indicate irrigation treatment. Asterisks indicate whether the difference between irrigated and non-irrigated is significant (p < 0.05 after Sidak adjustment for multiple comparisons) for the given genotype on the given planting date.

```{r, echo = FALSE, fig.height = 7}
# Arrange all genotypes with the same order
gen_order <- as.character(cld1_yld_genotype$Strains[order(cld1_yld_genotype$emmean)])

emmeans1_yld_pdxirrxgeno <- emmeans(lmm_yield_bystrain, ~ Irr | PD + Strains, mode = 'kenward-roger')
cld1_yld_pdxirrxgeno <- cld(emmeans1_yld_pdxirrxgeno, adjust = 'sidak', Letters = letters)

cld1_yld_pdxirrxgeno$Strains <- factor(cld1_yld_pdxirrxgeno$Strains, levels = gen_order)
cld1_yld_pdxirrxgeno$Irr <- factor(cld1_yld_pdxirrxgeno$Irr, levels = c('Irr', 'Nonirr'), labels = c('irrigated', 'non-irrigated'))

pd <- position_dodge(width = 0.5)
colors_irr <- scale_color_manual(name = '', values = c('slateblue', 'indianred'))

# Asterisks for each irrigation comparison
dat_cld <- copy(cld1_yld_pdxirrxgeno)
setDT(dat_cld)
signif_cld <- dat_cld[, .(signif = 'b' %in% trimws(.group), y_pos = max(upper.CL) + 500), by = .(PD, Strains)]
signif_cld[, symbol := ifelse(signif, '*', '')]

ggplot(cld1_yld_pdxirrxgeno, aes(x = Strains, y = emmean, label = gsub(' ', '', .group))) +
  geom_errorbar(width = 0.15, size = 0.75, aes(color = Irr, group = interaction(Strains, Irr), ymin = lower.CL, ymax = upper.CL), position = pd) +
  geom_point(aes(group = interaction(Strains, Irr)), size = 2, position = pd) + 
  geom_text(data = signif_cld, aes(label = symbol, y = y_pos), size = 8) +
  scale_y_continuous(name = parse(text = 'Yield~(kg~ha^-1)'), labels = scales::comma, limits = c(0, 5500)) +
  theme(legend.position = c(0.25, 0.95),
        legend.direction = 'horizontal',
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = 'Genotypes') +
  facet_wrap(~ PD, nrow = 3) + colors_irr
```

